### 2016-07-06 simulation/statfail.Rmd
slight improvements in graphics, hopefully clearer phrasing (that also matches new phrasing in paper)

---
title: 'A common misapplication of statistical inference: nuisance control with null-hypothesis significance test'
author: "Phillip M. Alday"
date: "May  2016"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
    self_contained: TRUE
    highlight: textmate
params:
  n: 500
  use_plotly: FALSE
  cores: 2
---

# Read the paper
Sassenhagen, Jona and Phillip M. Alday (under review): A common misapplication of statistical inference: nuisance control with null-hypothesis significance tests. Brain & Language. [arXiv preprint](http://arxiv.org/abs/1602.04565). [Repository](https://github.com/jona-sassenhagen/statfail/).

The title will probably change in the near feature, as some of our initial reviewer feedback has been some very helpful terminological suggestions.

<!--Another Python-based simulation will soon be made available on Binder, but the preliminary code is already in the GitHub repository.-->

A Shiny app that allows you to more closely individual simulated experiments for a given set of simulation parameters is available in the GitHub repository and on [shinyapps.io](https://palday.shinyapps.io/statfail/).
If you're going to do lots of computations, please run the app locally so that server time remains available for others.

# Simulation
Here, we repeat the simulation in the Shiny app, but iterate over different possibilities for

1. **manipulation effect size:** the effect size from the manipulation of interest (given as Cohen's $d$ between groups)
2. **confound size:** the measured size of the confounding feature (given as Cohen's $d$ between groups)
3. **confound-outcome correlation:** the simple correlation between that the confound size and the outcome (given as Pearson's $r$)

The last two emphasize a subtle point -- one of the many problems with doing inferential tests on group attributes (e.g. word frequency vs. condition in language studies) is that you're testing the difference in the feature and not the impact of that (difference in the) feature on the outcome. 
In other words, you're assuming that the measured feature difference exactly correlates with the impact that feature has on the outcome, which is a fairly strong assumption.

In other words, the confound-outcome correlation is the actual effect size of the confounding variable, taken together with the confound size, you can compute the actual impact of confounding on the manipulation.

## Setup

```{r setup, include=FALSE}
set.seed(42) # use a fixed random seed for reproducability

# use some parallel magic to speed things long
library(parallel)
options(mc.cores=params$cores)
library(doMC)
library(foreach)
registerDoMC(params$cores)

library(reshape2)
library(plyr)
library(ggplot2)
# if plotly isn't loaded, then just pass through ggplot's output
if(params$use_plotly){ 
  library(plotly) 
}else{
  ggplotly <- function(x) x
}
# this is the meat of the simulation
source("statfail.R")
```

The difficult part of the simulation is performed by `statfail.R`, which is shared with the Shiny app.
Both the source code and an explanation of the general principles (`simulation_notes.md`) are available in the GitHub repository.
Here, we only need to iterate over a few possibilities for the manipulation effect size, confound size, and confound-outcome  correlation.
The number of simulated experiments is held constant for all combinations of these (set via the document parameter `n`, which had the value `r params$n` when this document was generated).
Increasing this value will of course help detect rare phenomena, but will drastically increase time and memory needed.

```{r simulate, cache = TRUE,message=FALSE,dependson=params}
results <- data.frame()
for(mes in seq(0,2,by=0.5)){
  for(cfs in seq(0,2,by=0.5)){
    for(cfec in c(0, 0.1, 0.3, 0.5, 0.8, 1)){  
      input <- list(n.sims=params$n,
                manipulation.effect.size=mes,               # in Cohen's d
                confound.feature.size=cfs,                  # in Cohen's d
                confound.feature.effect.correlation = cfec, # Pearson
                n.items=20)
      simulation <- resimulate(n=input$n.sims
                      ,manipulation.effect.size=input$manipulation.effect.size
                      ,confound.feature.size=input$confound.feature.size
                      ,confound.feature.effect.correlation=input$confound.feature.effect.correlation
                      ,n.items=input$n.item
                      ,lapply.fnc=mclapply)

      pretest <- compute.feature.stats(simulation,.parallel=TRUE)
      manipulation <- compute.manipulation.regression(simulation,.parallel=TRUE)
      feature <- compute.feature.regression(simulation,.parallel=TRUE)
      multiple <- compute.multiple.regression(simulation,.parallel=TRUE)

      r <- compute.aggregate.results(pretest,manipulation,feature,multiple)
      cmp <- mclapply(r,function(x) x < 0.05)
      cmp <- as.data.frame(cmp)

      ## the rejections based on the pretest
      rejections <- sum(cmp$pretest) / input$n.sims
      manipulation_still_significant <- sum(with(subset(cmp,pretest), manipulation.multiple)) / input$n.sims
      feature_has_no_effect <- sum(with(subset(cmp,pretest), !feature.simple)) / input$n.sims
      feature_irrelevant_in_multiple <- sum(with(subset(cmp,pretest), !feature.multiple)) / input$n.sims

      ## the acceptances based on the pretest
      not_rejected <-  sum(!cmp$pretest) / input$n.sims
      feature_relevant_in_multiple <- sum(with(subset(cmp,!pretest), feature.multiple)) / input$n.sims
      manipulation_only_in_simple <-  sum(with(subset(cmp,!pretest), manipulation.simple & !manipulation.multiple)) / input$n.sims

      rsum <- data.frame(rejections
                         ,manipulation_still_significant
                         ,feature_has_no_effect
                         ,feature_irrelevant_in_multiple
                         ,not_rejected
                         ,feature_relevant_in_multiple
                         ,manipulation_only_in_simple
                         ,mes,cfs,cfec)

      results <- rbind(results, rsum)
    }
  }
}

# Convert the data to long format for easy plotting
rejections <- melt(results
                   ,id.vars=c("mes","cfs","cfec")
                   ,measure.vars=c("rejections"
                                   ,"manipulation_still_significant" 
                                   ,"feature_has_no_effect"
                                   ,"feature_irrelevant_in_multiple")
                   ,variable.name="type"
                   ,value.name = "percent.studies")
rejections$result <- "reject"

acceptances <- melt(results
                    ,id.vars=c("mes","cfs","cfec")
                    ,measure.vars=c("not_rejected"
                                    ,"feature_relevant_in_multiple"
                                    ,"manipulation_only_in_simple")
                    ,variable.name="type"
                    ,value.name = "percent.studies")
acceptances$result <- "accept"

results.long <- join(rejections,acceptances,type="full")
results.long$percent.studies <- results.long$percent.studies * 100
```

## Generate graphics
```{r generate.graphics, dependson='simulate',cache=TRUE,message=FALSE,warning=FALSE}
# prettier names for the legend
results.long$type <-factor(results.long$type
                           ,levels=c("rejections"
                                     ,"manipulation_still_significant"
                                     ,"feature_has_no_effect"
                                     ,"feature_irrelevant_in_multiple"
                                     ,"not_rejected"
                                     ,"feature_relevant_in_multiple"
                                     ,"manipulation_only_in_simple")
                           ,labels = c("rejected"
                                       ,"but significant manipulation missed"
                                       ,"when feature was not significant in simple regression"
                                       ,"when feature was not significant in multiple regression"
                                       ,"accepted"
                                       ,"but feature still had an effect in multiple regression",
                                       "and the manipulation was only significant in simple regression"))

# This uses the deprecated labeller API for ggplot.
facet_labeller <- function(variable,value){
  lbl <- NULL
  if (variable=='cfs') {
    lbl <- paste("Feature:",value)
  }else if (variable == 'mes') {
    lbl <- paste("Manipulation:",value)
  }else if (variable == 'cfec') {
    lbl <- paste("Feature-outcome correlation",value)
  } else {
    lbl <- as.character(value)
  }
}

# generate the graphic for rejections
gg.reject <- ggplot(subset(results.long,result=="reject"),aes(color=type,x=cfec,y=percent.studies)) +
  geom_line(stat="identity") + 
  facet_grid(cfs ~ mes,labeller = facet_labeller) + 
  xlab("Correlation of confounding feature and dependent variable") + 
  ylab("Percent of simulated studies") + 
  theme(legend.position="top", axis.text.x = element_text(angle = 45, hjust = 1), strip.text.y = element_text(angle=0)) + #coord_fixed(ratio=1/100) +
  guides(color=guide_legend(title="Result based on inferential pretesting of confounding feature",
                            title.position="top",
                            nrow=2))

# generate the same graphic for acceptances
gg.accept <- gg.reject %+% subset(results.long,result=="accept")

```

# Results  {.tabset .tabset-fade .tabset-pills}
The results in the following are whether or not a particular experimental manipulation was "allowed", i.e. accepted or rejected, based on inferential testing on the difference in confounding feature between experimental groups.

- *Manipulation* is the effect size in Cohen's $d$ from the experimental manipulation.
- *Feature* is the difference in Cohen's $d$ of the confounding feature between groups (again **not** its effect).

## Rejections
* Rejections where the significant manipulation was missed are an example of throwing the baby out with the bathwater.
* Rejections where the confounding feature was not a significant predictor in simple or multiple regression are arguably rejections where there is no effect of the feature.

```{r reject, echo=FALSE, fig.align="center", cache=FALSE}
ggplotly(gg.reject)
```

## Acceptances
* Acceptances where the feature still had an effect in multiple rejection are failed detections of a real confound.
* Acceptances where the manipulation was only significant in simple regression are failed detections of a real confound *that completely subsumes the manipulation effect*.

```{r accept, echo=FALSE, fig.align="center", cache=FALSE}
ggplotly(gg.accept)
```